On the Origin of Shrimpoluminescence

### 2022-01-14 sonoluminescence/paper/bak/v4/paper.tex
sdfadsfa

\emph{Any} ensemble of atoms at a well defined temperature will emit radiation according to Planck's law. In the case of SBSL, we will look for corrections due to interactions of the emitted photons with atoms on the way out of the bubble \cite{hilgenfeldt1999sonoluminescence}. Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we calculated before. Planck's law is \cite{schroeder1999introduction}
\begin{equation}
    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}
\end{equation}
The \emph{black-body model} assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out. Then the only emission that is measured comes from a thin layer on the bubbles surface. Calculations of SBSL assuming this over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}. Instead, we need to consider finite opacity of the bubble, i.e. some of the emitted radiation is absorbed on its way out.

Let's call the absorption coefficient $\kappa(\lambda,T(t))$\footnote{Really, this is an \emph{effective} absortion coeffecient $\kappa = \kappa'(1-\exp\left[ hc/k_B T \lambda \right])$ with $\kappa'$ the actual absoption coefficient of the material. The effective coefficient $\kappa$ includes induced emission of photons as light passes through the medium which 'decreases' absorption. Physically, this can be viewed as radiation that is absorbed being immediately re-emitted or, equivalently, vice-versa \cite{zel2002physics,taylor1969experimental}.} where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$. Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics}
\begin{equation}
    I(\lambda;T(t))=I_{B}(\lambda;T(t))  \left( 1-\exp \left[ -\kappa(\lambda;T(t)) s \right] \right)
    \label{eq:intensity}
\end{equation}
where have continued with the approximation that the temperature in the bubble is uniform \cite{hilgenfeldt1999sonoluminescence,hilgenfeldt1999simple}. In the limit that $\kappa\rightarrow\infty$, we recover the result for a black-body emitter. On the otherhand, for a strong absorber, $1-\exp(\kappa s)\approx\kappa s$ and $I\approx I_B \kappa s$, i.e. the number of photons emitted is greatly reduced, a result that we require. For more complicated $\kappa(\lambda;T(t))$ that may even depend on \emph{position} through $T(\bm{x},t)$ in a more accurate gas dynamical model, eq. \ref{eq:intensity} is integrated numerically. Moreover, integrating over the bubbles volume and surface, and over time, allow calculating the radiated power spectrum which may be directly compared to experiment. A great deal of work has been devoted to identifying the relevant absorption mechanisms, with most progress being made only after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.

What was discovered is that the radiative processes occurring during SBSL depend very sensitively on the fluid properties and driving pressure \cite{flannigan2006measurement,flannigan2005plasma}. Due to its different mechanical properties, H$_2$SO$_4$ can support SBSL over a larger range of driving pressures than water \cite{an2009diagnosing}. For low driving pressures, the temperature and pressure in the bubble are much lower than those encountered in bubbles in water and \emph{discrete} emission lines for molecular dissociation and electronic transitions are directly observable \cite{flannigan2005plasma,flannigan2006measurement}. At larger driving pressures leading to bubble conditions comparable to those in water, however, what was found is that the \emph{continous} processes electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination dominate \cite{hilgenfeldt1999simple,yasui2018acoustic,suslick2008inside,zel2002physics}. Moreover, it is now known that radiative (de)attachement of electrons and water-vapor molecules for SBSL in water is an important source of absorption, though this effect is limited by the small amount of these species in the bubble (see the argon rectification hypothesis above) \cite{an2006mechanism}.

@@

The \emph{Saha equation} of astrophysics allows us to estimate the fraction of particles that are ionized at a given temperature \cite{hilgenfeldt1999simple,an2006mechanism,zel2002physics}
\begin{equation}
    \alpha(T(t)) \approx \left( \frac{2 \pi m_e k_B T}{h^2} \right)^{3/4} \exp \left(-\frac{E_{ion}}{2 k_B T} \right)
\end{equation}
It turns out that at typical SBSL temperatures, only $\sim 1\%$ of the atoms are ionized and thus, since collisions are very frequent, electron-neutral atom bremsstrahlung is perhaps as important as electron-ion processes \cite{hilgenfeldt1999simple,an2006mechanism,flannigan2005plasma}. For Ar$^+$ and $e$, formulas for recombination and electron-ion bremsstrahlung can be found in ref. \cite{hilgenfeldt1999simple}. For electron-atom processes, most modern work follows that of Geltman \cite{geltman1973free,an2006mechanism,an2008spectral}.

For electron-ion collision, the potential leading to bremsstrahlung is the long range Coulumb interaction and analytical calculations are available. For electron-atom collision, the potential is short range and the emission is very sensitive to the potential. The best route is numerical calculations with pseudopotentials \cite{geltman1973free}.

Book \cite{zel2002physics} gives a very good heurestic explanation of the the absorptions processes in III.2. Also interesting: "Let us imagine an infinite medium in thermodynamic equilibrium at a constant temperature T. Under steady state conditions the radiation field will also be in equilibrium. Thermal radiative equilibrium is characterized by the fact that the number of photons or amount of radiant energy emitted by the medium per unit time per unit volume in a given frequency interval $d\nu$ and in a given differential solid angle $d\Omega$ is exactly equal to the number of absorbed photons or to the radiant energy absorbed by the medium in the same intervals $d\nu$ and $d\Omega$." Also "Kirchhoff's law expresses the general principle of detailed balancing applied to the emission and absorption of light. It allows us to calculate the emission coefficient of a substance if the absorption coefficient is known (and vice versa)."

Gives the three important processes: (i) free-free transitions such as electron-ion and electron-atom \emph{bremsstrahlung}. (ii) bound-free transitions such as detachment/recombination of electrons and ions, and (iii) bound-bound transitions such as transitions between bound electronic states in an atom. Both (i) and (ii) produce continous spectra, while (iii) leads to \emph{lines} \cite{zel2002physics}.

Early work ignored bound-bound transitions since lines weren't observed in spectrum. Argued that they would be very broad, but estimated the broadening was too small to lead to what was seen in experiment, concluding instead that these process don't occur and that the relevant species aren't present in the bubble \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}. Later work used more accurate gas dyanmical models to show which species are important \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of the spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}.

>>>>

\emph{Any} ensemble of atoms at a well defined temperature will emit radiation according to Planck's law. In the case of SBSL, we will look for corrections due to interactions of the emitted photons with atoms on the way out of the bubble \cite{hilgenfeldt1999sonoluminescence}. Since we are assuming LTE, the intensity of emitted radiation depends instantaneoisly on the temperature-time curve, $T(t)$, that we calculated before. Planck's law gives the energy radiated per unit-time and unit-volume, $I_B$, \cite{schroeder1999introduction}
\begin{equation}
    I_{B}(\lambda;T(t))=\frac{2 h c^2}{\lambda^5}\frac{1}{\exp(hc/\lambda k_B T(t))-1}.
\end{equation}
The subscript $B$ is to remind us this is for a \emph{black-body} which assumes the bubble is optically thick and that all radiation emitted inside the bubble is absorbed on its way out. Then the only emission that is measured comes from a thin layer on the bubble's surface. Calculations of SBSL assuming black-body radiation over-estimate the number of photons emitted by orders of magnitude \cite{hilgenfeldt1999simple,brenner2002single}. Instead, we need to consider finite opacity of the bubble.

Let's call the absorption coefficient $\kappa(\lambda,T(t))$\footnote{Really, this is an \emph{effective} absortion coeffecient $\kappa = \kappa'(1-\exp\left[ hc/k_B T \lambda \right])$ with $\kappa'$ the actual absoption coefficient of the material. The effective coefficient $\kappa$ includes induced emission of photons as light passes through the medium which 'decreases' absorption. Physically, this can be viewed as radiation that is absorbed being immediately re-emitted or, equivalently, vice-versa \cite{zel2002physics,taylor1969experimental}.} where we allow dependence on both wavelength $\lambda$ and temperature, $T(t)$. Then the intensity of light that has travelled a distance $s$ through the bubble is denoted by \cite{zel2002physics,hilgenfeldt1999simple}
\begin{equation}
    I(\lambda;T(t))=I_{B}(\lambda;T(t))  \left( 1-\exp \left[ -\kappa(\lambda;T(t)) s \right] \right)
    \label{eq:intensity}
\end{equation}
where have continued with the approximation that the temperature in the bubble is uniform \cite{hilgenfeldt1999sonoluminescence,hilgenfeldt1999simple}. In the limit that $\kappa\rightarrow\infty$, we recover the result for a black-body emitter. On the otherhand, for a strong absorber, $1-\exp(\kappa s)\approx\kappa s$ and $I\approx I_B \kappa s$, i.e. the number of photons emitted is greatly reduced, a result that we require. For more complicated $\kappa(\lambda;T(t))$ (that may even depend on \emph{position} through $T(\bm{x},t)$ in a more accurate gas dynamical model) eq. \ref{eq:intensity} is integrated numerically. Moreover, integrating $I(\lambda,T(t))$ over the bubbles and over time allows calculating the radiated power spectrum which may be directly compared to experiment. A great deal of work has been devoted to identifying the relevant absorption mechanisms leading to $\kappa$, with most progress being made only after SBSL was discovered in aqueous H$_2$SO$_4$ \cite{hilgenfeldt1999simple,brenner2002single,hilgenfeldt1999sonoluminescence,yasui1999mechanism,flannigan2005plasma,flannigan2006measurement,suslick2008inside,an2009diagnosing,an2008spectral,an2006mechanism}.

What is now known is that the radiative processes occurring during SBSL depend very sensitively on the fluid properties and driving pressure \cite{flannigan2006measurement,flannigan2005plasma}. Due to its different mechanical properties, H$_2$SO$_4$ can support SBSL over a larger range of driving pressures than water \cite{an2009diagnosing}. For low driving pressures, the temperature and pressure in the bubble are much lower than those encountered in bubbles in water and \emph{discrete} emission lines for molecular dissociation and electronic transitions are directly observable \cite{flannigan2005plasma,flannigan2006measurement}. At larger driving pressures leading to bubble conditions comparable to those in water, however, what was found is that the \emph{continous} processes electron-ion bremsstrahlung, electron-atom bremsstrahlung, and electron-ion recombination dominate \cite{hilgenfeldt1999simple,yasui2018acoustic,suslick2008inside,zel2002physics}. Moreover, it is now known that recombination of water-vapor molecules for SBSL in water is an important source of absorption, though this effect is limited by the small amount of these species in the bubble (see the argon rectification hypothesis above) \cite{an2006mechanism}.

@@

The \emph{Saha equation} of astrophysics allows us to estimate the fraction of particles that are ionized at a given temperature \cite{hilgenfeldt1999simple,an2006mechanism,zel2002physics}
\begin{equation}
    \frac{N_+ N_e}{N} \approx \left( \frac{2 \pi m_e k_B T}{h^2} \right)^{3/2} \exp \left(-\frac{E_{ion}}{ k_B T} \right)
    \label{eq:saha}
\end{equation}
It turns out that at typical SBSL temperatures, only $\sim 1\%$ of the atoms are ionized and thus, since collisions are very frequent, electron-neutral atom bremsstrahlung is perhaps as important as electron-ion processes \cite{hilgenfeldt1999simple,an2006mechanism,flannigan2005plasma}. For Ar$^+$ and $e$, formulas for recombination and electron-ion bremsstrahlung can be found in ref. \cite{hilgenfeldt1999simple}. For electron-atom processes, most modern work follows that of Geltman \cite{geltman1973free,an2006mechanism,an2008spectral}.

For electron-ion collision, the potential leading to bremsstrahlung is the long range Coulumb interaction. Below, we will derive a classical formula which is the most commonly used expressions in the context of SBSL. For electron-atom collisions, the potential is short range and the emission is very sensitive to the potential. The best route is numerical calculations with pseudopotentials \cite{geltman1973free}, though classical approximations similar to the one we will derive below have been used too \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence,zel2002physics}. Full quantum calculations of electron-ion bremsstrahlung absorption have also been done (see ref. for a review \cite{eckert2015aether}).

The instantaneous radiation energy emitted by an electron is\footnote{This is evaluated on the surface of a sphere at a very large distance from the electron; the time, $t$, should be the retarded time. But since we are going to look at spectral properties, we won't worry about this detail.} \cite{zel2002physics,griffiths2005introduction}
\begin{equation}
    S(t)=\frac{2e^2}{3c^2}\bm{a}^2(t)
\end{equation}
where $\bm{a}$ is the acceleration and $e$ the electron charge (we are using CGS units here). This can be deduced by calculating the Poynting vector and finding the terms that don't vanish at infinite distance. If we Fourier transform $\bm{a}(t) \rightarrow \bm{a}(\nu)$, this becomes
\begin{equation}
    S(\nu)=\frac{16 \pi^2 e^2}{3c^2}\bm{a}^2(\nu)
\end{equation}
with $\nu=c/\lambda$ the ordinary frequency.

Now consider a parallel beam of electrons with initial velocity $v$, and impact parameter $b$, incident from infinity with constant number density $N_e$. Integrating over the azimuthal angle, there are $v N_e \cdot  2\pi b db$ electrons passing per unit time through a ring with area $2\pi b db$ centered on the ion. Integrating over all possible orbits, the energy radiated \emph{per-ion} per frequency interval $d\nu$, for unit electron flux, $v N_e=1$, is the so-called ``effective radiation" $d\epsilon_\nu$ \cite{zel2002physics} 
\begin{equation}
    d\epsilon_\nu = d\nu \int_0^\infty S(\nu) 2 \pi b db.
    \label{eq:effective_radiation}
\end{equation}
This can be integrated for the motion of electron in the field of an ion with charge $Ze$ to give \cite{zel2002physics,landau2013classical}
\begin{equation}
    d\epsilon_\nu = \frac{32 \pi^2 }{3 \sqrt{3}} \frac{Z^2e^6}{m^2 c^3 v^2} d\nu
    \label{eq:effective_radiation_2}
\end{equation}

Now suppose that, in a unit volume of gas, there are $N_+$ ions with charge $Ze$ and $N_e$ electrons with velocities distributed according to $f(v)dv=4\pi(m/2\pi k_B T)^{3 / 2} \exp(-mv^2 / 2 k_B T) v^2 dv$, i.e. the \emph{Maxwell velocity distribution}, passing through the volume. The energy emitted per unit-volume and unit-time in the frequency interval $\nu+d\nu$ by the electrons moving throught the field of the ions is 
\begin{equation}
    N_+ N_e f(v') dv' v' d\epsilon_\nu(v').
    \label{eq:emission}
\end{equation}
Note, we are assuming the velocity of the ions is neglibigle and that there is no recoil. The energy emitted by all of the electrons is obtained by integrating from $v_{min}$ to $v=\infty$, with $v_{min}$ satisfying $1/2 m_e v^2_{min}=h\nu$. $v_{min}$ is the minimum velocity of an electron that is able to emit a photon. Using eq. \ref{eq:effective_radiation_2} and integrating, we find the emission coefficient, $j_\nu$, due to electron-ion bremsstrahlung:
\begin{equation}
    j_\nu d \nu=\frac{32\pi}{3}\left( \frac{2\pi}{3k_B T m_e} \right)^{1/2} \frac{Z^2 e^6}{m_e c^3} N_+ N_e \exp\left( -\frac{h\nu}{k_B T} \right)d\nu.
\end{equation}

This is a nice result, but we want the \emph{absorption coefficient}. Following nearly identical arguments, the energy in the interval $\nu+d\nu$ \emph{absorbed} per unit-volume and unit-time is
\begin{equation}
    N_+ N_e \cdot I_B(\nu;T) d\nu \cdot f(v) dv \cdot a_\nu \left[ 1-\exp\left( -\frac{h\nu}{k_B T} \right)\right].
    \label{eq:absorption}
\end{equation}
Here, $a_\nu$ is the absorption per-ion per-electron with velocity $v$ and the factor $1-\exp(-h\nu/k_B T)$ accounts for effective decrease in absorption due to stimulated emission \cite{zel2002physics}. In equilibrium, which we assume in SBSL, emission is exactly cancled by absorption. Then we set eq. \ref{eq:emission} equal to eq. \ref{eq:absorption} to solve for $a_\nu$. Integrating eq. \ref{eq:absorption} over $v$, we finally arrive at the spectral absorption coeffecient for a gas of electrons and ions with number densisites $N_e$ and $N_+$ respectively
\begin{equation}
    \kappa_\nu = \frac{4}{3} \left( \frac{2\pi}{3 m_e k_B T(t)} \right)^{1/2} \frac{Z^2 e^6}{h c m_e \nu^3} N_+ N_e .
\end{equation}
In terms of the wavelength, $\lambda$, and in SI units, we arrive at the equation used to calculate the SBSL spectrum in nearly all modern work \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing}
\begin{equation}
    \kappa_\lambda = \frac{4}{3} \left( \frac{2\pi}{3 m_e k_B T(t)} \right)^{1/2} \frac{Z^2 e^6 \lambda^3}{ (4 \pi \epsilon_0)^3 h c^4 m_e } N_+ N_e .
    \label{eq:absorption_coeff}
\end{equation}

We are finally have a complete model for SBSL! An RP equation, e.g. eq. \ref{eq:RP_3}, combined with a gas-dynamical equation, e.g. eq. \ref{eq:p_g}, can be used to solve for the dynamics of a bubble undergoing SBSL. This allow's us to calculate the temperature as a function of time, $T(t)$, according to e.g. eq. \ref{eq:T(t)}.  Then eq. \ref{eq:saha} can be inserted into \ref{eq:absorption_coeff} and, with expressions for electron-atom bremsstrahlung \cite{geltman1973free} and electron-ion recombination \cite{hilgenfeldt1999simple,an2006mechanism,an2008spectral,an2009diagnosing,zel2002physics}, eq. \ref{eq:intensity} can be integrated to calculate the power spectrum as a function of $T(t)$ and the power spectrum can be directly compared to experiment.

Early work on SBSL ignored bound-bound transitions since lines weren't observed in the SBSL spectrum in water. It was that they would be very broad due to the extremely high pressure in the bubble, but estimated broadening was too small to lead to what was seen in experiment and it was concluded instead that these process don't occur \cite{hilgenfeldt1999simple,hilgenfeldt1999sonoluminescence}. Later work using more accurate gas dyanmical models have shown that bound-bound transitions are important under the right conditions \cite{an2006mechanism,an2008spectral,an2006mechanism} and, when correctly including the broadening, pretty much all features of SBSL spectra in water and aqueous H$_2$SO$_4$ are predicted correclty \cite{suslick2008inside,flannigan2006measurement,flannigan2005plasma}. SBSL is a solved problem!
